Think deep for reef resilience

### 2025-08-15 paper/paper_draft.qmd
Move description of connectivity related aspect to same section
Also note comment for @eco-ben

@@

Connectivity data from 24 days in total were used in this investigation. These data span six
years: 2015, 2016, 2018, and 2019 are each represented by 3 daily connectivity matrices,
while 2017 and 2020 are each represented by 5 daily matrices. Indicative connectivity
strengths are implemented in ADRIA-CoralBlox as the mean of all available data for each
location (i.e., all 24 days). While the connectivity strengths do not vary from year to year
within the simulation, they do vary between reefs of varying sizes, providing sufficient
heterogeneity to capture spatial patterns in larval dispersal and settlement dynamics. The
spatial variability also enables investigations into the influence of connectivity strength
and its role in driving recovery processes.

@@

The relationship between the number of years above $\theta_{0}$ and median reef depth was
then assessed with Spearman's rank correlation. Spearman's correlation coefficient was
adopted as reef depth and connectivity strength distributions are expected to be non-normal
and highly skewed. Furthermore, reefs are naturally of different types and experience
different conditions, and so the monotonic trends (whether the relationship consistently
increase or decrease) rather than the exact quantative relationship was of interest for this
analysis.

Reefs in each timeseries cluster were compared to identify characteristics that differ
between low, medium and high coral cover clusters. These characteristics include reef median
depth, mean timeseries DHW, mean timeseries DHW – tolerance level, Pearson’s correlation
between total coral cover and inversed DHW timeseries, proportion of coral habitable area
occupied by corals in the first timestep, ratio of source-to-sink larval connectivity and
total strength of all larval connections between reefs. To analyse the DHW - tolerance
levels of reefs, heritable DHW – tolerance values were tracked over timeseries in a timestep
\* taxa \* reef \* scenario matrix. The mean timeseries DHW - tolerance was calculated over
the first four scenarios for each GCM (four from 4096 scenarios per GCM) due to
computational restraints, and then used to calculate an overall mean DHW – tolerance value
for each reef under each GCM.

The ratio of source-to-sink connectivity ($O:I$) for a target reef represents the strength of the reef's outgoing larval flow relative to
it's incoming larval flow (@eq-so-to-si). If a reef has no incoming connections, the $O:I$ equals the
$\frac{S_O}{N_O}$ value. Incoming connections in this calculation, do not directly represent
the larval supply gained as the population from which connections are coming from is not
represented. A source-to-sink ratio greater than 1 (a value greater than 0 on a log scale)
indicates a reef has greater outgoing than incoming larval flow.

$$
O:I = \frac{\frac{S_O}{N_O}}{\frac{S_I}{N_I}}
$$ {#eq-so-to-si}

-   $S$ denotes the total strength of connections.

-   $O$ denotes outgoing (source) connections from the target reef.

-   $I$ denotes incoming (sink) connections to the target reef.

>>>>

Connectivity data from 24 days in total were used in this investigation. These data span six
years from 2015 to 2020. The years 2015, 2016, 2018 and 2019 are each represented by
connectivity matrices covering spawning periods over three days in each year (12 days in
total), while 2017 and 2020 are each represented by five spawning days each, coming to a
total of 22 days of connectivity matrices. Indicative connectivity strengths are implemented
in ADRIA-CoralBlox as the mean of all available data for each location (i.e., mean of all 22
spawning days). While the connectivity strengths do not vary from year to year within each
simulation, they do vary between reefs whiuch are of varying sizes, providing sufficient
heterogeneity to capture spatial patterns in larval dispersal and settlement dynamics. The
spatial variability also enables investigations into the influence of connectivity strength
and its role in driving recovery processes.

The ratio of source-to-sink connectivity ($O:I$) for a target reef represents the strength
of the reef's outgoing larval flow relative to it's incoming larval flow (@eq-so-to-si). If
a reef has no incoming connections, the $O:I$ equals the $S_O/N_O$ value. Incoming
connections in this calculation, do not directly represent the larval supply gained as the
population from which connections are coming from is not represented. A source-to-sink ratio
greater than 1 (a value greater than 0 on a log scale) indicates a reef has greater outgoing
than incoming larval flow.

$$ O:I = \frac{(S_O/N_O)}{(S_I/N_I)} $$

-   $S$ is the sum of outgoing and incoming connections, indicating the total connectivity
    strength.

-   $O$ denotes outgoing (source) connections from a reef.

-   $I$ denotes incoming (sink) connections to a reef.

@@

Reefs in each timeseries cluster were compared to identify characteristics that differ
between low, medium and high coral cover clusters. These characteristics include reef median
depth, mean timeseries DHW, mean timeseries DHW tolerance level, Pearson’s correlation
between total coral cover and inversed DHW timeseries, proportion of coral habitable area
occupied by corals in the first timestep, ratio of source-to-sink larval connectivity and
total strength of all larval connections between reefs. To analyse the DHW tolerance levels
of reefs, heritable DHW tolerance values were tracked over timeseries in a matrix of shape
timestep \* taxa \* reef \* scenario matrix. The mean timeseries DHW tolerance was
calculated over the first four scenarios for each GCM (four from 4096 scenarios per GCM) due
to computational restraints, and then used to calculate an overall mean DHW tolerance value
for each reef under each GCM. **\[TI: Ben, is the last sentence here necessary?\]**

@@

The relationship between the number of years above $\theta_{0}$ and median reef depth was
then assessed with Spearman's rank correlation. Spearman's correlation coefficient was
adopted as reef depth and connectivity strength distributions are expected to be non-normal
and highly skewed. Furthermore, reefs are naturally of different types and experience
different conditions, and so the monotonic trends (whether the relationship consistently
increase or decrease) rather than the exact quantitative relationship was of interest for
this analysis.
